Potential of I/O Aware Workflows in Climate and Weather

### 2020-05-25 paper.tex
Incorporated Luciana's comments.

  \item \textbf{Script} executes the commands with one or multiple (potentially parallel) applications to run sequentially.
  During the creation of the script, Cylc has included variables that describe the task in the workflow.
  The information is typically fed into the application(s) representing the task and defining the storage location.
  The script uses commands to generate filenames considering the cycle and may store data in a workflow-specific shared directory.
  Either these commands are set in the Cylc workflow and then injected as environment variables or directly utilised as a part of the user-provided script.

@@

ESDM is used here as the I/O middleware in the parallel application (with NetCDF or directly) and orchestrates the I/O according to a simplified ESDM configuration file (\texttt{esdm.conf}).
This file contains information about the available technology in the data centre, its I/O characteristics, and will be used to make decisions about how to prioritise I/O targets.
In the example presented in \cref{lst:esdm.conf}, we have three storage targets: two global accessible file systems (\texttt{lustre01} and \texttt{lustre02}), and one local file system in \texttt{/tmp} that can be accessed via the POSIX backend.
Each of them comes with a lightweight performance model and the maximum size of data fragments.
The metadata section (Line 24) utilises here a POSIX interface to store the information about ESDM objects. Internally, ESDM creates so-called containers and dataset objects to manage data fragments.

@@

\cref{pic/data} shows the read/write performance when using ESDM to store a time series of 10 steps of a variable with 200k $\times$ 200k dimension (about 1\,km global resolution of the model, equivalent to 3\,TB of data in total).
Note that, while we only consider the volume data for one variable with one level and one timestep, this value could be multiplied by a sensible number of levels and timesteps of the model.
Data is stored on either Lustre02 or both Lustre file systems -- ESDM splits data of a single variable internally and distributes them across the file systems.
While the mapping is not yet optimal, the figure shows that the write performance benefits from this approach.
% Note that each Lustre file system has 128 Object Storage Targets (OSTs)
% (250 in total)
% which means that performance will not improve beyond 500 nodes.

>>>>

  \item \textbf{Script} starts the commands sequentially (a command can be a parallel application).
  During the creation of the script, Cylc has included variables that describe the task in the workflow.
  The information is typically fed into the application(s) representing the task and defining the storage location.
  The script uses commands to generate filenames considering the cycle and may store data in a workflow-specific shared directory.
  Either these commands are set in the Cylc workflow and then injected as environment variables or directly utilised as a part of the user-provided script.

@@

The system information of the design is realized using already available capabilities in the ESDM middleware.
We assume ESDM is used as the I/O middleware in the parallel application (with NetCDF or directly) and orchestrates the I/O according to a simplified ESDM configuration file (\texttt{esdm.conf}).
This file contains information about the available technology in the data centre, its I/O characteristics, and will be used to make decisions about how to prioritise I/O targets.
In the example presented in \cref{lst:esdm.conf}, we have three storage targets: two global accessible file systems (\texttt{lustre01} and \texttt{lustre02}), and one local file system in \texttt{/tmp} that can be accessed via the POSIX backend.
Each of them comes with a lightweight performance model and the maximum size of data fragments.
The metadata section (Line 24) utilises here a POSIX interface to store the information about ESDM objects. Internally, ESDM creates so-called containers and dataset objects to manage data fragments.

@@

\cref{pic/data} shows the read/write performance when using ESDM to store a time series of 10 steps of a variable with 200k $\times$ 200k dimension (about 1\,km global resolution of the model, equivalent to 3\,TB of data in total).
Note that, while we only consider the volume data for one variable with one level and ten timesteps, this value could be multiplied by a sensible number of levels and timesteps of the model.
Data is stored on either Lustre02 or both Lustre file systems -- ESDM splits data of a single variable internally and distributes them across the file systems.
While the mapping is not yet optimal, the figure shows that the write performance benefits from this approach.
In our observations, performance will not improve beyond 500 nodes, which might be due to the fact that each Lustre file system has 128 Object Storage Targets (OSTs) (250 in total).
